---
title: "Exploring Diamonds dataset"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    theme: simplex

---

```{r setup, include=FALSE}
library(shiny)
library(plotly) # devtools::install_github('ropensci/plotly')
library(ggplot2) # devtools::install_github('hadley/ggplot2')
library(dplyr)
library(flexdashboard) # devtools::install_github('rstudio/flexdashboard')
library(readr)
library(ggthemes)
library(DT) ##showing the dataset
library(devtools)
install_github("vqv/ggbiplot")
library(ggbiplot)
# Read in data
data("diamonds")
```

Sidebar {.sidebar}
======================================================================

### Controls for Univariate and Bivariate
```{r}
selectInput("uni", label = h4("Select a categorical variable"), 
                  choices = list("cut" = "cut", "clarity" = "clarity",
                                 "color" = "color"))
nms <- names(diamonds)
selectInput('v1', 'Variable1 for t-test', choices = nms, selected = "price")
selectInput('v2', 'Variable2 for t-test ', choices = nms, selected = "caret")
```


### Controls for Multivariate 

```{r}
nms <- names(diamonds)
selectInput('x', 'X', choices = nms, selected = "carat")
selectInput('y', 'Y', choices = nms, selected = "price")
selectInput('color', 'Color', choices = nms, selected = "clarity")
selectInput('facet_row', 'Facet Row', c(None = '.', nms))
selectInput('facet_col', 'Facet Column', c(None = '.', nms))
sliderInput('plotHeight', 'Height of plot (in pixels)', 
                    min = 100, max = 2000, value = 1000)
selectInput('com1', 'First Component', choices = list( "PC1" = "PC1", "PC2" = "PC2",
                                 "PC3" = "PC3", "PC4" = "PC4","PC5" = "PC5",
                                 "PC6" = "PC6","PC7" = "PC7"))
selectInput('com2', 'Second Component',choices = list( "PC2" = "PC2", "PC1" = "PC1",
                                 "PC3" = "PC3", "PC4" = "PC4","PC5" = "PC5",
                                 "PC6" = "PC6","PC7" = "PC7"))
checkboxInput('center', label='Center', value = TRUE)
checkboxInput('scale', label='Scale.', value = TRUE)
```


### Controls for price prediction
```{r}
helpText("This application predicts the price of a diamond based on its characteristics.")
h3(helpText("Select:"))
numericInput("car", label = h4("Carats"), step = 0.01, value = 2)
selectInput("cut", label = h4("Cut"), 
                  choices = list( "Fair" = "*", "Good" = "^Good",
                                 "Very Good" = "Very Good", "Premium" = "Premium",
                                 "Ideal" = "Ideal"))
selectInput("col", label = h4("Color"), 
                  choices = list("D" = "*", "E" = "E",
                                 "F" = "F", "G" ="G",
                                 "H" = "H", "I" = "I",
                                 "J" = "J"))
selectInput("clar", label = h4("Clarity"), 
                  choices = list("I1" = "*", "SI2" = "SI2",
                                 "SI1" = "SI1", "VS2" = "VS2", "VS1" = "VS1",
                                 "VVS2" = "VVS2", "VVS1" = "VVS1", "IF" = "IF" ))
```


click each tab to explore more about this project. Thank you very much~  

Author: Jishan Luo

Dataset
======================================================================

Row
-----------------------------------------------------------------------

### View the Diamonds dataset

```{r}
DT::renderDataTable({
    DT::datatable(diamonds, options = list(lengthMenu = c(10, 30, 50), pageLength = 5))
  })
```

Row 
-----------------------------------------------------------------------

### Know your data

#### The number of rows and columns
```{r}
dim(diamonds)
```
#### Variable names
```{r}
names(diamonds)
```

#### String of the dataset
```{r}
str(diamonds)#data frame by observations and variables
```

### The summary of the dataset

```{r}
summary(diamonds)
```


Univariate and Bivariate
============================================================================

Row
----------------------------------------------------------------------------

### Bar plot

```{r }
output$barplot <- renderPlotly({ 
  ggplot(diamonds, aes_string(x=input$uni, fill=input$uni)) + geom_bar( ) +
  scale_fill_brewer(palette = "Set1")
})

plotlyOutput('barplot')
```

Row
-----------------------------------------------------------------------------
```{r}
renderPrint({
var1=as.matrix(diamonds[,input$v1])
var2=as.matrix(diamonds[,input$v2])
t1=t.test(var1,var2)
t1
})
```


Row
-----------------------------------------------------------------------------
### Frequency Distribution table for cut

```{r}
renderPrint({
w1=table(diamonds$cut)
t1=as.data.frame(w1)
t1
})
```




### Frequency Distribution table for color

```{r}
renderPrint({
w1=table(diamonds$color)
t1=as.data.frame(w1)
t1
})

```

### Frequency Distribution table for clarity

```{r}
renderPrint({
w1=table(diamonds$clarity)
t1=as.data.frame(w1)
t1
})
```




Multivariate 1
============================================================================
Row
-----------------------------------------------------------------------------
```{r}
library(GGally)
ggcorr(diamonds)
```

Row
-----------------------------------------------------------------------------
```{r}
renderPlotly({

    # build graph with ggplot syntax
    p <- ggplot(diamonds, aes_string(x = input$x, y = input$y, color =diamonds$cut)) + 
      geom_point()

    # if at least one facet column/row is specified, add it
    facets <- paste(input$facet_row, '~', input$facet_col)
    if (facets != '. ~ .') p <- p + facet_grid(facets)

    ggplotly(p) %>% 
      layout(height = input$plotHeight, autosize=TRUE)

  })
```


Multivariate 2
============================================================================
Row
-----------------------------------------------------------------------------
### PCA plot
```{r}
diamonds.pca <- prcomp(diamonds[,c(1,5:10)], center = TRUE,scale. = TRUE)
renderPlotly({
diamonds.pca <- prcomp(diamonds[,c(1,5:10)], center = TRUE,scale. = TRUE)
ggbiplot(diamonds.pca,ellipse=TRUE, labels=rownames(diamonds), groups=diamonds$cut)
})
```


### PCA summary
```{r}
summary(diamonds.pca)
```


Row
-----------------------------------------------------------------------------
### PCA plot--Scatter plot with two components
```{r}
df_out <- as.data.frame(diamonds.pca$x)

renderPlotly({
PC1=df_out$PC1
PC2=df_out$PC2
PC3=df_out$PC3
PC4=df_out$PC4
PC5=df_out$PC5
PC6=df_out$PC6
PC7=df_out$PC7
p<-ggplot(df_out,aes(x=input$com1,y=input$com2,color=diamonds$cut))
p<-p+geom_point()
p
})
```

```{r eval=FALSE}
p<-ggplot(df_out,aes(x=PC1,y=PC2,color=diamonds$cut))
p<-p+geom_point()
p
```

### PCA Variance plots
```{r}
plot(diamonds.pca, type = "l")
```


Diamond price predictor
============================================================================

Row
-----------------------------------------------------------------------------
### Relationship plot among variables using ggplot
```{r}
output$distPlot <- renderPlotly({
    # select diamonds depending of user input
    diam <- filter(diamonds, grepl(input$cut, cut), grepl(input$col, color), grepl(input$clar, clarity))
    # build linear regression model
    fit <- lm( price~carat, diam)
    # predicts the price 
    pred <- predict(fit, newdata = data.frame(carat = input$car,
                                              cut = input$cut,
                                              color = input$col,
                                              clarity = input$clar))
    # Drow the plot using ggplot2
    plot <- ggplot(data=diam, aes(x=carat, y = price))+
      geom_point(aes(color = cut), alpha = 0.3)+
      geom_smooth(method = "lm")+
      geom_vline(xintercept = input$car, color = "red")+
      geom_hline(yintercept = pred, color = "green")+ 
      ggtitle("ggplot for diamonds")
    plot
  })
  
  
  
 plotlyOutput('distPlot') 
 
```

Row
----------------------------------------------------------------------------

```{r}
#The plot for relationship among the variables of diamonds
output$result <- renderText({
    # renders the text for the prediction below the graph
    diam <- filter(diamonds, grepl(input$cut, cut), grepl(input$col, color), grepl(input$clar, clarity))
    fit <- lm( price~carat, diam)
    pred <- predict(fit, newdata = data.frame(carat = input$car,
                                              cut = input$cut,
                                              color = input$col,
                                              clarity = input$clar))
    res <- paste("$",round(pred, digits = 2))
    print(res)
  })

h4("Predicted value of this diamond is:")
h3(textOutput("result"))
```



About
============================================================================

For this project I used the Diamonds dataset from package `ggplot2` in R studio. After loading the dataset, I decided to answer the the following question: What is the basic summmary of this dataset? How can I utilize the statistics knowledge to analyze this dataset? What is the predicted price of a diamond based on the attributes of cut, color, clarity and carat? What is the accuracy of the prediction? Additionally, I decided to to answer additional questions in order to learn some additional methods / processes and code that may be of use in the future. 

First, let's have a look at this dataset

```{r,echo=TRUE}
library(plotly)
library(ggplot2)
library(readr)
library(tidyr)
library(dplyr)
library(GGally)
library(rpart.plot)
require(memisc)
data("diamonds") #read diamonds dataset
head(diamonds)
str(diamonds)#data frame by observations and variables
dim(diamonds)
names(diamonds)
summary(diamonds)
sum(is.na(diamonds)) # check the values to see if I have any NAs in my data and it doesn't seem to indicate that I do. 
```

Second, I used univarite, bivariate, and multivariate methods to analyze the dataset.Also I used plots to demonstrate the result I want to present.



